---
title: "Basic Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Population adjustment methods such as matching-adjusted indirect comparison (MAIC) are increasingly used to compare marginal treatment effects when there are cross-trial differences in effect modifiers and limited patient-level data. MAIC is based on propensity score weighting, which is sensitive to poor covariate over- lap and cannot extrapolate beyond the observed covariate space. Current out- come regression-based alternatives can extrapolate but target a conditional treatment effect that is incompatible in the indirect comparison. When adjusting for covariates, one must integrate or average the conditional estimate over the rel- evant population to recover a compatible marginal treatment effect. We propose a marginalization method based on parametric G-computation that can be easily applied where the outcome regression is a generalized linear model or a Cox model. The approach views the covariate adjustment regression as a nuisance model and separates its estimation from the evaluation of the marginal treatment effect of interest. The method can accommodate a Bayesian statistical framework, which naturally integrates the analysis into a probabilistic framework. A simula- tion study provides proof-of-principle and benchmarks the method's performance against MAIC and the conventional outcome regression. Parametric G- computation achieves more precise and more accurate estimates than MAIC, par- ticularly when covariate overlap is poor, and yields unbiased marginal treatment effect estimates under no failures of assumptions. Furthermore, the marginalized regression-adjusted estimates provide greater precision and accuracy than the conditional estimates produced by the conventional outcome regression, which are

## Example

```{r setup, warning=FALSE, message=FALSE}
library(boot)      # non-parametric bootstrap in MAIC and ML G-computation
library(copula)    # simulating BC covariates from Gaussian copula
library(rstanarm)  # fit outcome regression, draw outcomes in Bayesian G-computation
library(mimR)
```

Implementing for MAIC, STC, G-computation methods on a simulated example.

```{r load-data}
set.seed(555)

AC.IPD <- read.csv(here::here("data", "AC_IPD.csv"))  # AC patient-level data
BC.ALD <- read.csv(here::here("data", "BC_ALD.csv"))  # BC aggregate-level data
```

This data consists of the following.

`AC.IPD`

-   `X*`: patient measurements
-   `trt`: treatment ID
-   `y`: indicator of whether event was observed

`BC.ALD`

-   `mean.X*`: mean
-   `sd.X*`: standard deviation
-   `y.*.sum`: total number of events
-   `N.*`: total number of individuals

```{r}
head(AC.IPD)
BC.ALD
```

We will create some function for the marginal variance, defined as

$$
\frac{1}{n_C} + \frac{1}{n_{\bar{C}}} + \frac{1}{n_B} + \frac{1}{n_{\bar{B}}}
$$

and the marginal treatment effect, defined as

$$
\log(n_B n_{\bar{C}}) - log(n_C n_{\bar{B}})
$$


## MAIC

Using the individual level data firstly perform non-parametric bootstrap
of the `maic.boot` function with `R` = 1000 replicates. This function
fits treatment coefficient for the marginal effect for $A$ vs $C$.
The returned value is an object of class `boot`. Then calculate the bootstrap mean and variance
in the wrapper function `maic_boot_stats`.

```{r hat_Delta_stats_maic}
hat_Delta_stats_maic <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_maic())
hat_Delta_stats_maic
```

Construct Wald-type normal distribution-based 95\% confidence interval.

## STC (conventional outcome regression)

Fit regression model of outcome on treatment and covariates
IPD effect modifiers centred at the mean BC values
purely prognostic variables are included but not centred.

```{r hat_Delta_stats_stc}
hat_Delta_stats_stc <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_stc())
hat_Delta_stats_stc
```

Next we perform G-computation firstly with a frequentist MLE approach and then a Bayesian approach.

## Parametric G-computation with maximum-likelihood estimation 

Matrix of pairwise correlations between IPD covariates.

```{r hat_Delta_stats_gcomp_ml}
# hat_Delta_stats_gcomp_ml <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_gcomp_ml())
```


## Bayesian G-computation with MCMC 

Matrix of pairwise correlations between IPD covariates.

```{r hat_Delta_stats_gcomp_stan}
hat_Delta_stats_gcomp_stan <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_gcomp_stan())
hat_Delta_stats_gcomp_stan
```

