---
title: "Comparison with other packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
  editor_options: 
  chunk_output_type: console
  chunk_output_type: console
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(outstandR)
```

The aim of this document is to compare the results from various other R packages with those produced using `{outstandR}`.
There are already a few widely-used R packages on CRAN to perform model-based standardization, such as [`{marginaleffects}`](https://marginaleffects.com/) and [`{stdReg}/{stdReg2}`](https://sachsmc.github.io/stdReg2/).

## marginaleffects

The `{marginaleffects}` package has functions to calculate G-computation estimates.

```{r , warning=FALSE, message=FALSE}
##TODO: see unit tests
library(marginaleffects)

# marginaleffects::avg_predictions()
# marginaleffects::avg_comparisons()
```


## stdReg[2]

The package `{stdReg}` has been rewritten to produce `{stdReg2}` which has "nicer output, more available methods, the possibility to include new methods, and mainly to make maintenance and updating easier." So, for this reason, we will use `{stdReg2}`.

We shall follow the article [_Estimation of causal effects using stdReg2_](https://sachsmc.github.io/stdReg2/articles/overview.html).
This is for continuous outcomes (change in weight) and includes transformed covariates.
Note that the regression has transformed covariates using the `I()` function. Two outcome scales are calculated, risk difference and risk ratio.
The data also contains categorical variable with more than 2 levels, `education`, `smokeintensity`, `exercise` and `active`.

```{r stdReg2-original, warning=FALSE, message=FALSE}
library(stdReg2)
library(causaldata)  # contains data set

# load data set
nhefs_dat <- causaldata::nhefs_complete

m <- glm(wt82_71 ~ qsmk + sex + race + age + I(age^2) + as.factor(education) + smokeintensity + I(smokeintensity^2) +
           smokeyrs + I(smokeyrs^2) + as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2), 
         data = nhefs_dat)
summary(m)

res_stdReg2 <- standardize_glm(
  wt82_71 ~ qsmk + sex + race + age + I(age^2) + as.factor(education) + smokeintensity + I(smokeintensity^2) +
    smokeyrs + I(smokeyrs^2) + as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2),
  data = nhefs_dat, 
  values = list(qsmk = c(0,1)),
  contrasts = c("difference", "ratio"),
  reference = 0)

res_stdReg2
```

Most simply, we can drop the variables with more than 2 levels and run `outstandR()`.

```{r stdReg2-alternative1, warning=FALSE, message=FALSE}
library(dplyr)

nhefs_ipd <- nhefs_dat |> 
  # drop categorical variables
  dplyr::select(qsmk, sex, race, age, smokeintensity, smokeyrs, wt71, wt82_71) |>
  dplyr::rename(trt = qsmk,
                y = wt82_71) |> 
  dplyr::mutate(sex = as.numeric(sex) - 1,
                race = as.numeric(race) - 1)

# create aggregate data

nhefs.X <- nhefs_ipd %>%
  summarise(across(-c(trt, y),
                   list(mean = mean, sd = sd),
                   .names = "{fn}.{col}"))

nhefs.B <- nhefs_ipd |> 
  dplyr::filter(trt == 1) |> 
  summarise(y.B.sum = sum(y),
            y.B.bar = mean(y),
            y.B.sd = sd(y),
            N.B = n())

nhefs.C <- nhefs_ipd |> 
  dplyr::filter(trt == 0) |> 
  summarise(y.C.sum = sum(y),
            y.C.bar = mean(y),
            y.C.sd = sd(y),
            N.C = n())

nhefs_ald <- cbind.data.frame(nhefs.X, nhefs.C, nhefs.B)

# simplified formula
lin_form <-
  as.formula(y ~ trt * (sex + race + age + I(age^2) +
                          smokeintensity + I(smokeintensity^2) +
                          smokeyrs + I(smokeyrs^2) + wt71 + I(wt71^2)))
nhefs_strat <- 
  strategy_gcomp_ml(formula = lin_form,
                    family = gaussian(link = "identity"))

res <- outstandR(BC.ALD = nhefs_ald,
                 AC.IPD = nhefs_ipd,
                 strategy = nhefs_strat,
                 scale = "log_relative_risk")
res
```

Alternatively, rather than dropping factors, we can create dummy variables to replace them using `maicplus::dummize_ipd()`.

```{r stdReg2-alternative2, warning=FALSE, message=FALSE}

nhefs_ipd <- nhefs_dat |> 
  dplyr::select(qsmk, sex, race, age, smokeintensity, smokeyrs, wt71,
                wt82_71, education, exercise, active) |> 
  dplyr::rename(trt = qsmk,
                y = wt82_71) |> 
  dplyr::mutate(sex = as.numeric(sex) - 1,       # binary
                race = as.numeric(race) - 1) |> 
  maicplus::dummize_ipd(dummize_cols = c("education", "exercise", "active"),
                        dummize_ref_level = c("1","0","0")) |> 
  dplyr::select(-education, -exercise, -active)  # remove originals

# # replace with this?
# model.matrix()

# create aggregate data

nhefs.X <- nhefs_ipd %>%
  summarise(across(-c(trt, y),
                   list(mean = mean, sd = sd),
                   .names = "{fn}.{col}"))

nhefs.B <- dplyr::filter(nhefs_ipd, trt == 1) |> 
  summarise(y.B.sum = sum(y),
            y.B.bar = mean(y),
            y.B.sd = sd(y),
            N.B = n())

nhefs.C <- dplyr::filter(nhefs_ipd, trt == 0) |> 
  summarise(y.C.sum = sum(y),
            y.C.bar = mean(y),
            y.C.sd = sd(y),
            N.C = n())

nhefs_ald <- cbind.data.frame(nhefs.X, nhefs.C, nhefs.B)

lin_form <-
  as.formula(y ~ trt * (sex + race + age + I(age^2) +
                          smokeintensity + I(smokeintensity^2) +
                          EDUCATION_2 + EDUCATION_3 + EDUCATION_4 + EDUCATION_5 +
                          EXERCISE_1 + EXERCISE_2 + ACTIVE_1 + ACTIVE_2 +
                          smokeyrs + I(smokeyrs^2) + wt71 + I(wt71^2)))
nhefs_strat <- 
  strategy_gcomp_ml(formula = lin_form,
                    family = gaussian(link = "identity"))

res <- outstandR(BC.ALD = nhefs_ald,
                 AC.IPD = nhefs_ipd,
                 strategy = nhefs_strat,
                 scale = "risk_difference")
res
```


---

The 2024 [ISPOR poster by Hatswell](https://www.ispor.org/heor-resources/presentations-database/presentation/euro2024-4016/141209) presents a comparison of different MAIC R packages. We shall use this as our starting point. Of the packages considered, `{MAIC}` is not on CRAN so not included in our analysis at the moment.

## maicplus

Roche developed the `maic` package which is now deprecated and replaced by `maicplus`.
We will follow the article titled [Anchored Binary Analysis](https://hta-pharma.github.io/maicplus/main/articles/anchored_binary.html) which is part of their package to make our comparison.

```{r maicplus-original, warning=FALSE, message=FALSE}
library(maicplus)

# Load required data
data(centered_ipd_twt)
data(adrs_twt)  # response IPD
data(adsl_sat)  # covariates IPD

# Load required package
library(dplyr)

# Prepare matching variables (only using means, ignoring SDs)
adsl_sat <- adsl_sat |> 
  mutate(
    SEX_MALE = as.integer(SEX == "Male"),
    AGE_SQUARED = AGE^2
  )

# Define variable names for matching (excluding SDs)
adsl_colnames <- c("AGE", "AGE_SQUARED", "SEX_MALE", "ECOG0", "SMOKE", "N_PR_THER_MEDIAN")
centered_colnames <- paste0(adsl_colnames, "_CENTERED")

# Estimate weights based only on means
weighted_data <- estimate_weights(
  data = centered_ipd_twt,
  centered_colnames = centered_colnames
)

# Aggregate summary data for AGD (ONLY MEANS, NO SDs)
agd <- data.frame(
  AGE_MEAN = 51,
  SEX_MALE_PROP = 147 / 300,
  ECOG0_PROP = 0.40,
  SMOKE_PROP = 58 / (300 - 5),
  N_PR_THER_MEDIAN = 2
)

# Binary AGD summary data
binary_agd <- data.frame(
  ARM = rep(c("B", "C"), each = 2),
  RESPONSE = rep(c("YES", "NO"), 2),
  COUNT = c(280, 200, 120, 200)
)

# Generate pseudo-IPD from binary AGD
pseudo_adrs <- get_pseudo_ipd_binary(
  binary_agd = binary_agd,
  format = "stacked"
)

# Perform MAIC analysis (matching only on means)
res_maicplus <- maic_anchored(
  weights_object = weighted_data,
  ipd = adrs_twt,
  pseudo_ipd = pseudo_adrs,
  trt_ipd = "A",
  trt_agd = "B",
  trt_common = "C",
  normalize_weight = FALSE,
  endpoint_type = "binary",
  endpoint_name = "Binary Endpoint",
  eff_measure = "OR",
  binary_robust_cov_type = "HC3"
)


res_maicplus
```

```{r maicplus-alternative, warning=FALSE, message=FALSE}
library(dplyr, verbose = FALSE, quietly = TRUE)
library(glue, verbose = FALSE, quietly = TRUE)

##TODO: why has this got median in name here?
adsl_colnames[adsl_colnames == "N_PR_THER_MEDIAN"] <- "N_PR_THER"

lin_form <- as.formula(glue("y ~ trt * ({paste(adsl_colnames, collapse = ' + ')})"))

AC.IPD <- adsl_twt |>
  merge(adrs_twt) |> 
  rename(trt = ARM,
         y = RESPONSE) |> 
  mutate(AGE_SQUARED = AGE^2)

# sd from IPD
agd_sd <- AC.IPD |> 
  summarise(across(c(AGE, AGE_SQUARED, SMOKE, ECOG0, N_PR_THER, SEX_MALE),
                   .fns = sd, .names = "sd.{.col}"))

# transform names to format needed in outstandR
# mean.* and y.*
# note that we're nto going to use sd in matching
BC.ALD <- agd |> 
  rename(mean.AGE = AGE_MEAN,
         # sd.AGE = AGE_SD,
         mean.SEX_MALE = SEX_MALE_PROP,
         mean.ECOG0 = ECOG0_PROP,
         mean.SMOKE = SMOKE_PROP,
         mean.N_PR_THER = N_PR_THER_MEDIAN) |> 
  mutate(
    # mean.AGE_SQUARED = sd.AGE^2 + mean.AGE^2,
    # mean.AGE_SQUARED = mean.AGE^2,                 ##TODO: this is a fudge
    # sd.SEX_MALE = sqrt(mean.SEX_MALE * (1 - mean.SEX_MALE)), 
    # sd.ECOG0 = sqrt(mean.ECOG0 * (1 - mean.ECOG0)), 
    # sd.SMOKE = sqrt(mean.SMOKE * (1 - mean.SMOKE)), 
    # outcomes
    N.B = 480, 
    y.B.sum = binary_agd[binary_agd$ARM == "B" & binary_agd$RESPONSE == "YES", "COUNT"],
    y.B.bar = y.B.sum/N.B,
    N.C = 320, 
    y.C.sum = binary_agd[binary_agd$ARM == "C" & binary_agd$RESPONSE == "YES", "COUNT"],
    y.C.bar = y.C.sum/N.C) |> 
    cbind(agd_sd) |> 
  relocate(N.B, y.B.sum, y.B.bar, N.C, y.C.sum, y.C.bar, .after = sd.SEX_MALE)

BC.ALD
```

Now that we've prepared the input data we're ready to run the model.

```{r eval=FALSE}
##TODO: fix bug of large value in optim()
outstandR_maic <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_maic(formula = lin_form,
                                     family = binomial(link = "logit")))
```


## maicChecks

We now follow the examples for the package [`maicChecks`](https://clyau.github.io/maicChecks/).

```{r, warning=FALSE, message=FALSE}
##TODO:
library(maicChecks)
```

---

For G-formula we will compare with the output from the packages `gFormulaMI` and `gfoRmula`.

## gfoRmula

```{r, warning=FALSE, message=FALSE}
##TODO:
library(gfoRmula)
```

## gFormulaMI

```{r, warning=FALSE, message=FALSE}
##TODO:
library(gFormulaMI)
```

