---
title: "Comparison with other packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comparison-with-other-packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(outstandR)
```

There are already a few widely-used R packages in CRAN to perform model-based standardization, such as `{marginaleffects}` and `{stdReg}`.

## marginaleffects

```{r , warning=FALSE, message=FALSE}
library(marginaleffects)
```


## stdReg

```{r, warning=FALSE, message=FALSE}
library(stdReg)
```

[ISPOR poster by Hatswell](https://www.ispor.org/heor-resources/presentations-database/presentation/euro2024-4016/141209) make a comparison of different MAIC R packages.

## maicplus

`MAIC` not on CRAN so not included at the moment. `maicChecks`

Roche developed `maic` which is now deprecated and replaced by `maicplus`.

For g-formula we will use `gFormulaMI` and `gfoRmula`.

## gfoRmula

```{r, warning=FALSE, message=FALSE}
library(gfoRmula)
```

## gFormulaMI

```{r, warning=FALSE, message=FALSE}
library(gFormulaMI)
```

