---
title: "Comparison with other packages"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{comparison-with-other-packages}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
  editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(outstandR)
```

There are already a few widely-used R packages in CRAN to perform model-based standardization, such as [`{marginaleffects}`](https://marginaleffects.com/) and [`{stdReg}/{StdReg2}`](https://sachsmc.github.io/stdReg2/).

## marginaleffects

```{r , warning=FALSE, message=FALSE}
library(marginaleffects)
```


## stdReg[2]

`{StdReg}` has been rewritten to produce `{StdReg2}` which has "nicer output, more available methods, the possibility to include new methods, and mainly to make maintenance and updating easier." So, for this reason, we will use `{StdReg2}`.

We shall follow the article [_Estimation of causal effects using stdReg2_](https://sachsmc.github.io/stdReg2/articles/overview.html)

```{r stdReg2-original, warning=FALSE, message=FALSE}
library(stdReg2)
library(causaldata)

# load dataset
# data(nhefs_complete, package = "causaldata")
nhefs_dat <- causaldata::nhefs_complete

m <- glm(wt82_71 ~ qsmk + sex + race + age + I(age^2) + as.factor(education) + smokeintensity + I(smokeintensity^2) + smokeyrs + I(smokeyrs^2) + as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2), 
         data = nhefs_dat)
summary(m)

m2 <- standardize_glm(
  wt82_71 ~ qsmk + sex + race + age + I(age^2) + as.factor(education) + smokeintensity + I(smokeintensity^2) + smokeyrs + I(smokeyrs^2) + as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2),
  data = nhefs_dat, 
  values = list(qsmk = c(0,1)),
  contrasts = c("difference", "ratio"),
  reference = 0)

m2
```

```{r stdReg2-alternative1, warning=FALSE, message=FALSE}
library(dplyr)

# factor of more than 2 levels
# education, exercise, active
# for now just remove them

##TODO: check how {outstandR} handles factor vs not factors?

##TODO: doubly-robust, separate treatment models?

##TODO: add explicit reference argument in set-up?
# reference = 0

nhefs_ipd <- nhefs_dat |> 
  select(qsmk, sex, race, age, smokeintensity, smokeyrs, wt71, wt82_71) |>
  rename(trt = qsmk,
         y = wt82_71) |> 
  mutate(sex = as.numeric(sex) - 1,
         race = as.numeric(race) - 1)

# create aggregate data

nhefs.X <- nhefs_ipd %>%
  summarise(across(-c(trt, y),
                   list(mean = mean, sd = sd),
                   .names = "{fn}.{col}"))

nhefs.B <- dplyr::filter(nhefs_ipd, trt == 1) |> 
  summarise(y.B.sum = sum(y),
            y.B.bar = mean(y),
            y.B.sd = sd(y),
            N.B = n())

nhefs.C <- dplyr::filter(nhefs_ipd, trt == 0) |> 
  summarise(y.C.sum = sum(y),
            y.C.bar = mean(y),
            y.C.sd = sd(y),
            N.C = n())

nhefs_ald <- cbind.data.frame(nhefs.X, nhefs.C, nhefs.B)

lin_form <-
  as.formula(y ~ trt * (sex + race + age + I(age^2) +
                          smokeintensity + I(smokeintensity^2) +
                          smokeyrs + I(smokeyrs^2) + wt71 + I(wt71^2)))

nhefs_strat <- 
  strategy_gcomp_ml(formula = lin_form,
                    family = gaussian(link = "identity"))

res <- outstandR(BC.ALD = nhefs_ald,
                 AC.IPD = nhefs_ipd,
                 strategy = nhefs_strat,
                 scale = "log_relative_risk")
res
```

Rather than dropping factors, create dummy variables.

```{r stdReg2-alternative2, warning=FALSE, message=FALSE}
nhefs_ipd <- nhefs_dat |> 
  select(qsmk, sex, race, age, smokeintensity, smokeyrs, wt71,
         wt82_71, education, exercise, active) |> 
  rename(trt = qsmk,
         y = wt82_71) |> 
  mutate(sex = as.numeric(sex) - 1,
         race = as.numeric(race) - 1) |> 
  maicplus::dummize_ipd(dummize_cols = c("education", "exercise", "active"),
                        dummize_ref_level = c("1","0","0")) |> 
  select(-education, -exercise, -active)

# create aggregate data

nhefs.X <- nhefs_ipd %>%
  summarise(across(-c(trt, y),
                   list(mean = mean, sd = sd),
                   .names = "{fn}.{col}"))

nhefs.B <- dplyr::filter(nhefs_ipd, trt == 1) |> 
  summarise(y.B.sum = sum(y),
            y.B.bar = mean(y),
            y.B.sd = sd(y),
            N.B = n())

nhefs.C <- dplyr::filter(nhefs_ipd, trt == 0) |> 
  summarise(y.C.sum = sum(y),
            y.C.bar = mean(y),
            y.C.sd = sd(y),
            N.C = n())

nhefs_ald <- cbind.data.frame(nhefs.X, nhefs.C, nhefs.B)

lin_form <-
  as.formula(y ~ trt * (sex + race + age + I(age^2) +
                          smokeintensity + I(smokeintensity^2) +
                          EDUCATION_2 + EDUCATION_3 + EDUCATION_4 + EDUCATION_5 +
                          EXERCISE_1 + EXERCISE_2 + ACTIVE_1 + ACTIVE_2 +
                          smokeyrs + I(smokeyrs^2) + wt71 + I(wt71^2)))
nhefs_strat <- 
  strategy_gcomp_ml(formula = lin_form,
                    family = gaussian(link = "identity"))

res <- outstandR(BC.ALD = nhefs_ald,
                 AC.IPD = nhefs_ipd,
                 strategy = nhefs_strat,
                 scale = "log_relative_risk")
res
```


---

[ISPOR poster by Hatswell](https://www.ispor.org/heor-resources/presentations-database/presentation/euro2024-4016/141209) make a comparison of different MAIC R packages.

Of the packages considered, `{MAIC}` is not on CRAN so not included in our anaysis at the moment.

## maicplus

Roche developed `maic` package which is now deprecated and replaced by `maicplus`.
We will follow the article titled [Anchored Binary Analysis](https://hta-pharma.github.io/maicplus/main/articles/anchored_binary.html).

```{r maicplus-original, warning=FALSE, message=FALSE}
library(maicplus)

data(centered_ipd_twt)
data(adrs_twt)

# original ipd
data("adsl_sat")

# Data containing the matching variables
adsl_sat <- adsl_sat |> 
  mutate(SEX_MALE = ifelse(SEX == "Male", 1, 0)) |> 
  mutate(AGE_SQUARED = AGE^2)

# # could use built-in function for dummizing variables
# adsl_sat <- dummize_ipd(adsl_sat,
#                         dummize_cols = "SEX",
#                         dummize_ref_level = c("Female"))

adsl_colnames <- c("AGE", "AGE_SQUARED", "SEX_MALE", "ECOG0", "SMOKE", "N_PR_THER_MEDIAN")
centered_colnames <- paste0(adsl_colnames, "_CENTERED")

weighted_data <- estimate_weights(
  data = centered_ipd_twt,
  centered_colnames = centered_colnames)

agd <- data.frame(
  AGE_MEAN = 51,
  AGE_SD = 3.25,
  SEX_MALE_PROP = 147 / 300,
  ECOG0_PROP = 0.40,
  SMOKE_PROP = 58 / (300 - 5),
  N_PR_THER_MEDIAN = 2
)

binary_agd <- data.frame(
  ARM = c("B", "C", "B", "C"),
  RESPONSE = c("YES", "YES", "NO", "NO"),
  COUNT = c(280, 120, 200, 200)
)

# get dummy binary IPD
pseudo_adrs <- get_pseudo_ipd_binary(
  binary_agd = binary_agd,
  format = "stacked"
)

result <- maic_anchored(
  weights_object = weighted_data,
  ipd = adrs_twt,
  pseudo_ipd = pseudo_adrs,
  trt_ipd = "A",
  trt_agd = "B",
  trt_common = "C",
  normalize_weight = FALSE,
  endpoint_type = "binary",
  endpoint_name = "Binary Endpoint",
  eff_measure = "OR",
  # binary specific args
  binary_robust_cov_type = "HC3"
)

result
```

```{r maicplus-alternative, warning=FALSE, message=FALSE}
library(dplyr)
library(glue)

# why has this got median in name here?
adsl_colnames[adsl_colnames == "N_PR_THER_MEDIAN"] <- "N_PR_THER"

lin_form <- as.formula(glue("y ~ trt * ({paste(adsl_colnames, collapse = ' + ')})"))

AC.IPD <- adsl_twt |>
  merge(adrs_twt) |> 
  rename(trt = ARM,
         y = RESPONSE) |> 
  mutate(AGE_SQUARED = AGE^2)

BC.ALD <- agd |> 
  rename(mean.AGE = AGE_MEAN,
         sd.AGE = AGE_SD,
         mean.SEX_MALE = SEX_MALE_PROP,
         mean.ECOG0 = ECOG0_PROP,
         mean.SMOKE = SMOKE_PROP,
         mean.N_PR_THER = N_PR_THER_MEDIAN) |> 
  mutate(mean.AGE_SQUARED = sd.AGE^2 + mean.AGE^2,
         sd.SEX_MALE = sqrt(mean.SEX_MALE * (1 - mean.SEX_MALE)), 
         sd.ECOG0 = sqrt(mean.ECOG0 * (1 - mean.ECOG0)), 
         sd.SMOKE = sqrt(mean.SMOKE * (1 - mean.SMOKE)), 
         # outcomes
         N.B = 480, 
         y.B.sum = binary_agd[binary_agd$ARM == "B" & binary_agd$RESPONSE == "YES", "COUNT"],
         y.B.bar = y.B.sum/N.B,
         N.C = 320, 
         y.C.sum = binary_agd[binary_agd$ARM == "C" & binary_agd$RESPONSE == "YES", "COUNT"],
         y.C.bar = y.C.sum/N.C)

##TODO: haven't got all the agd sd. do I need it?
outstandR_maic <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_maic(formula = lin_form,
                                     family = binomial(link = "logit")))
```


## maicChecks

We now follow the examples for [`maicChecks`](https://clyau.github.io/maicChecks/).

```{r, warning=FALSE, message=FALSE}
library(maicChecks)
```

---

For g-formula we will use `gFormulaMI` and `gfoRmula`.

## gfoRmula

```{r, warning=FALSE, message=FALSE}
library(gfoRmula)
```

## gFormulaMI

```{r, warning=FALSE, message=FALSE}
library(gFormulaMI)
```

