---
title: "COPD Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Analysis

First, let us load necessary packages.

```{r setup, warning=FALSE, message=FALSE}
library(boot)      # non-parametric bootstrap in MAIC and ML G-computation
library(copula)    # simulating BC covariates from Gaussian copula
library(rstanarm)  # fit outcome regression, draw outcomes in Bayesian G-computation
library(outstandR)
```

### Data

```{r load-data}
set.seed(555)

AC.IPD <- read.csv(here::here("raw-data", "AC_IPD.csv"))  # AC patient-level data
BC.ALD <- read.csv(here::here("raw-data", "BC_ALD.csv"))  # BC aggregate-level data
```

This general format of data sets consist of the following.

#### `AC.IPD`: Individual patient data

-   `X*`: patient measurements
-   `trt`: treatment ID
-   `y`: (logical) indicator of whether event was observed

#### `BC.ALD`: Aggregate-level data

-   `mean.X*`: mean patient measurement
-   `sd.X*`: standard deviation of patient measurement
-   `y.*.sum`: total number of events
-   `y.*.bar`: total number of events
-   `N.*`: total number of individuals

Note that the wildcard `*` here is usually an integer from 1 or the trial identifier _B_, _C_.

Our data look like the following.

```{r}
head(AC.IPD)
```

There are 4 correlated continuous covariates generated per subject, simulated from a multivariate normal distribution.

```{r}
BC.ALD
```

### Output statistics



## Model fitting in R


### MAIC

The formula used in this model is

$$
y = X_3 + X_4 + \beta_t X_1 + \beta_t X_2
$$
which corresponds to the following `R` `formula` object passed as an argument to the strategy function.

```{r}
lin_form <- as.formula("y ~ X3 + X4 + trt*X1 + trt*X2")
```


```{r outstandR_maic}
outstandR_maic <- outstandR(AC.IPD, BC.ALD, strategy = strategy_maic(formula = lin_form))
```

The returned object is of class `outstandR`.

```{r outstandR_maic-print}
outstandR_maic
```


### STC

$$
g(\mu_n) = \beta_0 + (\boldsymbol{x}_n - \boldsymbol{\theta}) \beta_1 + (\beta_z + (\boldsymbol{x_n^{EM}} - \boldsymbol{\theta^{EM}}) \boldsymbol{\beta_2}) \; \mbox{I}(z_n=1)
$$

$$
y = X_3 + X_4 + \beta_t(X_1 - \bar{X_1}) + \beta_t(X_2 - \bar{X_2})
$$
However, `outstandR()` knows how to handle this so we can simply pass the same (uncentred) formula as before.

```{r outstandR_stc}
outstandR_stc <- outstandR(AC.IPD, BC.ALD, strategy = strategy_stc(formula = lin_form))
outstandR_stc
```


### Parametric G-computation with maximum-likelihood estimation 


$$
g(\mu_n) = \beta_0 + \boldsymbol{x}_n \boldsymbol{\beta_1} + (\beta_z + \boldsymbol{x_n^{EM}} \boldsymbol{\beta_2}) \; \mbox{I}(z_n = 1)
$$


$$
\hat{\mu}_0 = \int_{x^*} g^{-1} (\hat{\beta}_0 + x^* \hat{\beta}_1 ) p(x^*) \; \text{d}x^*
$$

$$
\hat{\Delta}^{(2)}_{10} = g(\hat{\mu}_1) - g(\hat{\mu}_0)
$$


```{r outstandR_gcomp_ml}
outstandR_gcomp_ml <- outstandR(AC.IPD, BC.ALD, strategy = strategy_gcomp_ml(formula = lin_form))
outstandR_gcomp_ml
```


### Bayesian G-computation with MCMC 


$$
p(y^*_{^z*} \mid \mathcal{D}_{AC}) = \int_{x^*} p(y^* \mid z^*, x^*, \mathcal{D}_{AC}) p(x^* \mid \mathcal{D}_{AC})\; \text{d}x^*
$$
$$
= \int_{x^*} \int_{\beta} p(y^* \mid z^*, x^*, \beta) p(x^* \mid \beta) p(\beta \mid \mathcal{D}_{AC})\; d\beta \; \text{d}x^*
$$

```{r outstandR_gcomp_stan}
outstandR_gcomp_stan <- outstandR(AC.IPD, BC.ALD, strategy = strategy_gcomp_stan(formula = lin_form))
outstandR_gcomp_stan
```

### Multiple imputation marginalisation


```{r outstandR_mim}
outstandR_mim <- outstandR(AC.IPD, BC.ALD, strategy = strategy_mim(formula = lin_form))
outstandR_mim
```

Combine all outputs

```{r}
knitr::kable(
  data.frame(
  `MAIC` = unlist(outstandR_maic$contrasts),
  `STC` = unlist(outstandR_stc$contrasts),
  `Gcomp ML` = unlist(outstandR_gcomp_ml$contrasts),
  `Gcomp Bayes` = unlist(outstandR_gcomp_stan$contrasts),
  `MIM` = unlist(outstandR_mim$contrasts))
)
```
