---
title: "Count data example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Count data example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

## Example analysis

First, let us load necessary packages.

```{r setup, warning=FALSE, message=FALSE}
library(boot)      # non-parametric bootstrap in MAIC and ML G-computation
library(copula)    # simulating BC covariates from Gaussian copula
library(rstanarm)  # fit outcome regression, draw outcomes in Bayesian G-computation
library(outstandR)
```

### Data

First, use the in-built `gen_data()` function.

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(MASS)
```

```{r}
N <- 200
allocation <- 2/3      # active treatment vs. placebo allocation ratio (2:1)
b_trt <- log(0.17)     # conditional effect of active treatment vs. common comparator
b_X <- -log(0.5)       # conditional effect of each prognostic variable
b_EM <- -log(0.67)     # conditional interaction effect of each effect modifier
meanX_AC <- c(0.45, 0.45)       # mean of normally-distributed covariate in AC trial
meanX_BC <- c(0.6, 0.6)         # mean of each normally-distributed covariate in BC
meanX_EM_AC <- c(0.45, 0.45)    # mean of normally-distributed EM covariate in AC trial
meanX_EM_BC <- c(0.6, 0.6)      # mean of each normally-distributed EM covariate in BC
sdX <- c(0.4, 0.4)     # standard deviation of each covariate (same for AC and BC)
sdX_EM <- c(0.4, 0.4)  # standard deviation of each EM covariate
corX <- 0.2            # covariate correlation coefficient  
b_0 <- -0.6            # baseline intercept coefficient  ##TODO: fixed value

AC.IPD <- gen_data(N, b_trt, b_X, b_EM, b_0,
                   meanX_AC, sdX, 
                   meanX_EM_AC, sdX_EM, 
                   corX, allocation,
                   family = poisson(link = "log"))
```

Similarly, for the aggregate data but with the additional summarise step.

```{r generate-ald-data}
BC.IPD <- gen_data(N, b_trt, b_X, b_EM, b_0,
                   meanX_BC, sdX, 
                   meanX_EM_BC, sdX_EM, 
                   corX, allocation,
                   family = poisson(link = "log"))

cov.X <- BC.IPD %>%
  summarise(across(starts_with("X"),
                   list(mean = mean, sd = sd),
                   .names = "{fn}.{col}"))

out.C <- dplyr::filter(BC.IPD, trt == 1) %>%
  summarise(y.B.sum = sum(y),
            y.B.bar = mean(y),
            N.B = n())

out.B <- dplyr::filter(BC.IPD, trt == 0) %>%
  summarise(y.C.sum = sum(y),
            y.C.bar = mean(y),
            N.C = n())

BC.ALD <- cbind.data.frame(cov.X, out.C, out.B)
```


### Output statistics

We will implement for MAIC, STC, and G-computation methods to obtain the
*marginal variance*, defined as

$$
$$

and the *marginal treatment effect*, defined as the log-odds ratio,

$$
$$

where $\bar{C}$ is the compliment of $C$ so e.g.
$n_{\bar{C}} = N_C - n_c$.

## Model fitting in R

The `{outstandR}` package has been written to be easy to use and
essential consists of a single function, `outstandR()`. This can be used
to run all of the different types of model, which we will call
*strategies*. The first two arguments of `outstandR()` are the
individual and aggregate-level data, respectively.

A `strategy` argument of `outstandR` takes functions called
`strategy_*()`, where the wildcard `*` is replaced by the name of the
particular method required, e.g. `strategy_maic()` for MAIC. Each
specific example is provided below.

### MAIC

Using the individual level data for *AC* firstly we perform
non-parametric bootstrap of the `maic.boot` function with `R = 1000`
replicates. This function fits treatment coefficient for the marginal
effect for *A* vs *C*. The returned value is an object of class `boot`
from the `{boot}` package. We then calculate the bootstrap mean and
variance in the wrapper function `maic_boot_stats`.

The formula used in this model is

$$
y = X_3 + X_4 + \beta_t X_1 + \beta_t X_2
$$

which corresponds to the following `R` `formula` object passed as an
argument to the strategy function.

```{r}
lin_form <- as.formula("y ~ X3 + X4 + trt*X1 + trt*X2")
```

```{r outstandR_maic}
outstandR_maic <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_maic(formula = lin_form,
                                     family = poisson(link = "log")))
```

The returned object is of class `outstandR`.

```{r outstandR_maic-print}
outstandR_maic
```

We see that this is a list object with 3 parts, each containing
statistics between each pair of treatments. These are the mean
contrasts, variances and confidence intervals (CI), respectively. The
default CI is for 95% but can be altered in `outstandR` with the `CI`
argument.

### STC

STC is the conventional outcome regression method. It involves fitting a
regression model of outcome on treatment and covariates to the IPD. IPD
effect modifiers are centred at the mean *BC* values.

$$
g(\mu_n) = \beta_0 + (\boldsymbol{x}_n - \boldsymbol{\theta}) \beta_1 + (\beta_z + (\boldsymbol{x_n^{EM}} - \boldsymbol{\theta^{EM}}) \boldsymbol{\beta_2}) \; \mbox{I}(z_n=1)
$$

where $\beta_0$ is the intercept, $\beta_1$ are the covariate
coefficients, $\beta_z$ and $\beta_2$ are the effect modifier
coefficients, $z_n$ are the indicator variables of effect alternative
treatment. $g(\cdot)$ is the link function e.g. $\log$.

As already mentioned, running the STC analysis is almost identical to
the previous analysis but we now use the `strategy_stc()` strategy
function instead and a formula with centered covariates.

$$
y = X_3 + X_4 + \beta_t(X_1 - \bar{X_1}) + \beta_t(X_2 - \bar{X_2})
$$

However, `outstandR()` knows how to handle this so we can simply pass
the same (uncentred) formula as before.

```{r outstandR_stc}
outstandR_stc <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_stc(formula = lin_form,
                                    family = poisson(link = "log")))
outstandR_stc
```

For the last two approaches, we perform G-computation firstly with a
frequentist MLE approach and then a Bayesian approach.

### Parametric G-computation with maximum-likelihood estimation

G-computation marginalizes the conditional estimates by separating the
regression modelling from the estimation of the marginal treatment
effect for *A* versus *C*. First, a regression model of the observed
outcome $y$ on the covariates $x$ and treatment $z$ is fitted to the
*AC* IPD:

$$
g(\mu_n) = \beta_0 + \boldsymbol{x}_n \boldsymbol{\beta_1} + (\beta_z + \boldsymbol{x_n^{EM}} \boldsymbol{\beta_2}) \; \mbox{I}(z_n = 1)
$$

In the context of G-computation, this regression model is often called
the “Q-model.” Having fitted the Q-model, the regression coefficients
are treated as nuisance parameters. The parameters are applied to the
simulated covariates $x*$ to predict hypothetical outcomes for each
subject under both possible treatments. Namely, a pair of predicted
outcomes, also called potential outcomes, under *A* and under *C*, is
generated for each subject.

By plugging treatment *C* into the regression fit for every simulated
observation, we predict the marginal outcome mean in the hypothetical
scenario in which all units are under treatment *C*:

$$
\hat{\mu}_0 = \int_{x^*} g^{-1} (\hat{\beta}_0 + x^* \hat{\beta}_1 ) p(x^*) \; \text{d}x^*
$$

To estimate the marginal or population-average treatment effect for *A*
versus *C* in the linear predictor scale, one back-transforms to this
scale the average predictions, taken over all subjects on the natural
outcome scale, and calculates the difference between the average linear
predictions:

$$
\hat{\Delta}^{(2)}_{10} = g(\hat{\mu}_1) - g(\hat{\mu}_0)
$$

```{r outstandR_gcomp_ml}
outstandR_gcomp_ml <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_gcomp_ml(formula = lin_form,
                                         family = poisson(link = "log")))
outstandR_gcomp_ml
```

### Bayesian G-computation with MCMC

The difference between Bayesian G-computation and its maximum-likelihood
counterpart is in the estimated distribution of the predicted outcomes.
The Bayesian approach also marginalizes, integrates or standardizes over
the joint posterior distribution of the conditional nuisance parameters
of the outcome regression, as well as the joint covariate distribution.

Draw a vector of size $N^*$ of predicted outcomes $y^*_z$ under each set
intervention $z^* \in \{0, 1\}$ from its posterior predictive
distribution under the specific treatment. This is defined as
$p(y^*_{z^*} \mid \mathcal{D}_{AC}) = \int_{\beta} p(y^*_{z^*} \mid \beta) p(\beta \mid \mathcal{D}_{AC}) d\beta$
where $p(\beta \mid \mathcal{D}_{AC})$ is the posterior distribution of
the outcome regression coefficients $\beta$, which encode the
predictor-outcome relationships observed in the *AC* trial IPD. This is
given by:

$$
p(y^*_{^z*} \mid \mathcal{D}_{AC}) = \int_{x^*} p(y^* \mid z^*, x^*, \mathcal{D}_{AC}) p(x^* \mid \mathcal{D}_{AC})\; \text{d}x^*
$$

$$
= \int_{x^*} \int_{\beta} p(y^* \mid z^*, x^*, \beta) p(x^* \mid \beta) p(\beta \mid \mathcal{D}_{AC})\; d\beta \; \text{d}x^*
$$

In practice, the integrals above can be approximated numerically, using
full Bayesian estimation via Markov chain Monte Carlo (MCMC) sampling.

The average, variance and interval estimates of the marginal treatment
effect can be derived empirically from draws of the posterior density.

We can draw a vector of size $N^*$ of predicted outcomes $y^*_z$ under
each set intervention $z^*$ from its posterior predictive distribution
under the specific treatment.

```{r outstandR_gcomp_stan}
outstandR_gcomp_stan <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_gcomp_stan(formula = lin_form,
                                           family = poisson(link = "log")))
outstandR_gcomp_stan
```

### Multiple imputation marginalisation

ref

$$
equation here
$$

```{r outstandR_mim}
outstandR_mim <-
  outstandR(AC.IPD, BC.ALD,
            strategy = strategy_mim(formula = lin_form,
                                    family = poisson(link = "log")))
outstandR_mim
```

### Model comparison

Combine all outputs for log-odds ratio table of all contrasts and methods.

```{r}
knitr::kable(
  data.frame(
  `MAIC` = unlist(outstandR_maic$contrasts),
  `STC` = unlist(outstandR_stc$contrasts),
  `Gcomp ML` = unlist(outstandR_gcomp_ml$contrasts),
  `Gcomp Bayes` = unlist(outstandR_gcomp_stan$contrasts),
  `MIM` = unlist(outstandR_mim$contrasts))
)
```
