---
title: "Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

```{r setup}
library(boot)      # non-parametric bootstrap in MAIC and ML G-computation
library(copula)    # simulating BC covariates from Gaussian copula
library(rstanarm)  # fit outcome regression, draw outcomes in Bayesian G-computation
library(mimR)
```

Implementing for MAIC, STC, G-computation methods on a simulated example.

```{r load-data}
set.seed(555)

AC.IPD <- read.csv(here::here("data", "AC_IPD.csv"))  # AC patient-level data
BC.ALD <- read.csv(here::here("data", "BC_ALD.csv"))  # BC aggregate-level data
```

This data consists of the following.

`AC.IPD`

-   `X*`: patient measurements
-   `trt`: treatment ID
-   `y`: indicator of whether event was observed

`BC.ALD`

-   `mean.X*`: mean
-   `sd.X*`: standard deviation
-   `y.*.sum`: total number of events
-   `N.*`: total number of individuals

```{r}
head(AC.IPD)
BC.ALD
```

We will create some function for the marginal variance, defined as

$$
\frac{1}{n_C} + \frac{1}{n_{\bar{C}}} + \frac{1}{n_B} + \frac{1}{n_{\bar{B}}}
$$

and the marginal treatment effect, defined as

$$
\log(n_B n_{\bar{C}}) - log(n_C n_{\bar{B}})
$$


## MAIC

Using the individual level data firstly perform non-parametric bootstrap
of the `maic.boot` function with `R` = 1000 replicates. This function
fits treatment coefficient for the marginal effect for $A$ vs $C$.
The returned value is an object of class `boot`. Then calculate the bootstrap mean and variance
in the wrapper function `maic_boot_stats`.

```{r}
hat_Delta_stats_maic <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_maic())
```

Construct Wald-type normal distribution-based 95\% confidence interval.

## STC (conventional outcome regression)

Fit regression model of outcome on treatment and covariates
IPD effect modifiers centred at the mean BC values
purely prognostic variables are included but not centred.

```{r}
hat_Delta_stats_stc <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_stc())
```

Next we perform G-computation firstly with a frequentist MLE approach and then a Bayesian approach.

## Parametric G-computation with maximum-likelihood estimation 

Matrix of pairwise correlations between IPD covariates.

```{r}
hat_Delta_stats_gcomp_ml <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_gcomp_ml())
```


## Bayesian G-computation with MCMC 

Matrix of pairwise correlations between IPD covariates.

```{r}
hat_Delta_stats_gcomp_stan <- hat_Delta_stats(AC.IPD, BC.ALD, strategy = strategy_gcomp_stan())
```

